Denotational Semantics of General Payment Primitives, and Its Payment System

### 2022-09-21 packages/spec-haskell/yellowpaper/Paper.tex
money distribution models

%% Import Packages
\usepackage{fvextra}
\usepackage{csquotes}
\usepackage{xcolor}
\usepackage{minted}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{graphicx}

@@

In Chapter I, we shall first explore the foundation for the survey. Here we present a formal definition of payment system,
and various models used in the system. We then select a few relevant approaches used in computer science useful for
modeling and defining formal specification for the payment system.

One of the approaches is \textit{denotational semantics}, which is used in Chapter II of the paper to define the general
payment primitives. Along with the denotational semantics, a restatement\footnote{It is a borrowed term from common law:
"restatement of the law". In our case, the denotative mathematical "laws".} of it in \textit{Haskell programming
    language} (\cite{hudak1992report} \cite{jones2003haskell} \cite{marlow2010haskell}) is also included.

In Chapter III, a reference implementation of the general payment primitives and its payment system called
\textit{Superfluid Money} is introduced.

In Chapter IV, some possible further investigations are included for future study purpose.

@@

We shall define a payment system, whereby:

@@

@@

Most of the existing money schemes are of this category, where a money unit completely disregards its context and hence
its liquidity value and bearer are static until the next money redistribution occurs.

Here are the type definitions of these money schemes defined in \cite{buldas2021unifying}:

\begin{figure}
  \includegraphics[width=\linewidth]{assets/discrete-money-distribution.png}
  \caption{Schematic representation of discrete money distribution}
  \label{fig:discrete-md}
\end{figure}

>>>>

%% Import Packages
\usepackage{fvextra}
\usepackage{csquotes}
\usepackage{xcolor}
\usepackage{minted}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{graphicx}

@@

In Chapter 2, we shall first explore the foundation for the survey. Here we present a formal definition of payment system,
and various models used in the system. We then select a few relevant approaches used in computer science useful for
modeling and defining formal specification for the payment system.

One of the approaches is \textit{denotational semantics}, which is used in Chapter 3 of the paper to define the general
payment primitives. Along with the denotational semantics, a restatement\footnote{It is a borrowed term from common law:
"restatement of the law". In our case, the denotative mathematical "laws".} of it in \textit{Haskell programming
    language} (\cite{hudak1992report} \cite{jones2003haskell} \cite{marlow2010haskell}) is also included.

In Chapter 4, a reference implementation of the general payment primitives and its payment system called
\textit{Superfluid Money} is introduced.

In Chapter 5, some possible further investigations are included for future study purpose.

@@

We first define a payment system, whereby:

@@

For its modern upgrade, the system should also have these properties:

\begin{itemize}
\item Money can be distributed continuously over time, as opposed to be discrete.
\item Payment primitives can involve more than two parties, as opposed to be only for a sender and a receiver.
\item Financial system should be compositional.
\end{itemize}

@@

A representation of money and its distribution proposed in \cite{buldas2021unifying} involves the following components:

\begin{itemize}
\item U is the set of monetary units
\item $\nu : U \rightarrow \mathbb{N}$ is the value function defining the value $\nu(u)$ of every value unit u. The set
    $\mathbb{N}$ is the set of all natural numbers, but instead, we can use any set of numerals that is totally ordered
    (e.g.\ integers, real numbers).
\item $\beta : U \rightarrow \mathfrak{B}$ is the bearer function defining the bearer $\beta(u)$ of a unit. The set
    $\mathfrak{B}$ is the set of possible bearers. The bearer is usually a legal construction defining any type of legal
    entity, such as a person, a family, a company, a state institution, etc.
\end{itemize}

This discrete nature of money distribution is schematically depicted in {\ref{fig:discrete-md}}:

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{../assets/discrete-money-distribution.png}
    \caption{Schematic representation of discrete money distribution}
    \label{fig:discrete-md}
\end{figure}
